---
title: "Analyzing Finch Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

Loading the packages necessary for the sript to run

```{r}
library(readxl)     # load readxl, for reading Excel files
library(tidyverse)  # load tidyverse, for working with datasets
```

## Read Finch Data

Reading the Finch Dataset

```{r}
finches <- read_excel("finches_data.xlsx")
```

## See all variables in the dataset

Glimpse the Finch Data Set

```{r}
glimpse(finches)
```

## Histogram

Creating histogram for beaklength, grouped by survival with labels

```{r}
ggplot(
  data = finches,                     # use the finches dataset
  mapping = aes(x = beak_length,      # put beak length on the x axis
                fill = outcome)       # fill sets the color of the boxes
) +
  geom_histogram(bins = 14) +         # add the histogram, use 14 bins
  facet_wrap(~ outcome, ncol = 1) +   # outcome is the grouping variable
  guides(fill = FALSE) +              # don't show a legend for fll color
  labs(
    title = "Figure 1.",              # title
    x = "Beak Length (mm)",           # x-axis label
    y = "Number of Birds"             # y-axis label
  )
```

## T-test

Perform for birds that died

```{r}
beak_length_died <-
  finches %>%                     # start with finches dataset
  filter(outcome == "died") %>%   # only include rows w/ outcome=died
  pull(beak_length)               # extract the beak_length column
```

Print vector

```{r}
beak_length_died
```

Perform for birds that died

```{r}
beak_length_survived <-
  finches %>% 
  filter(outcome == "survived") %>% 
  pull(beak_length)
```

Print vector

```{r}
beak_length_survived
```

Perform a two-tailed t-test on survived versus died

```{r}
t.test(beak_length_died, beak_length_survived)
```

